---
title: "Script 3: Clean / Processs NESARC Dataset"
author: "Angel Garcia de la Garza"
date: "March 25th, 2022"
output: md_document
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, 
                      results = 'asis',
                      cache = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      cache.lazy = FALSE)

library(pROC)
library(summarytools)
library(doParallel)
library(beepr)
library(summarytools)
library(psych)
library(knitr)
library(forcats)
library(furrr)
library(randomForest)
library(gbm)
library(e1071)
library(caret)
library(glmnet)
library(tidyverse)



```

This script processes and organizes the risk factors variables in the data. This script retypes variables into categorical vs. numeric, identifies duplicated variables, and codes missing values. A more detailed explanation of this process is included in the supplement (see "Organizing the risk factor variables").

# Load Data

```{r load data, cache = FALSE}

load(here("Data","Raw_Data", "w1w2suicidedata.RData"))
load(here("Data","Raw_Data", "w1w2suicidedata_WithCases_May2020.RData"))

```


# Create Suicide Variables

```{r}

w1w2suicide <- w1w2suicide.v3

rm(w1w2suicide.v3)

w1w2suicide <- w1w2suicide %>%
                mutate(suicide_sli = suicide_new) %>%
                mutate(W2S4AQ3A16 = ifelse(is.na(W2S4AQ3A16), 
                                           2, W2S4AQ3A16),
                       suicide_334 = ifelse(W2S4AQ3A16 == 1,
                                            1, suicide_sli)) %>%
                rename(suicide_222 = suicide_sli) %>%
                select(-W2S4AQ3A16,
                       -W2S14Q17A,
                       -W2S14Q17B,
                       -W2S14Q17C,
                       -suicide_life,
                       -suicide_age,
                       -age_diff,
                       -suicide_new,
                       -W2AGE)

```



# Sort out numeric from non numeric

I created a decision tree that uses two flag the numeric variables. 

* All those variables that include number in either it's name or description
* All those variables that have more than 20 unique values

## Find those variables in the code book with "number"

```{r clean_values}

w1vars1 <- w1vars1 %>%
            mutate(num.description = str_detect(description, "NUM"),
                   scale.description = str_detect(description, "SCALE"),
                   age.description = str_detect(description, "AGE AT"),
                   num.name = str_detect(code, "NUM"),
                   num.logical = num.description | scale.description | num.name | age.description)

```

## Find variables with more than 20 unique values

```{r}

unique_values <- function(x) {
  length(unique(x))
}

w1w2_datatype <- w1w2suicide %>% 
                  map_dfc(~unique_values(.)) %>%
                  gather(key = "code", "unique") %>%
                  mutate(cont_twenty_logic = unique > 19)

```

## Merge these two together to create one single rule

```{r}

w1w2_datatype <- w1w2_datatype %>%
                  left_join(w1vars1, by = "code") %>%
                  mutate(is_numeric = cont_twenty_logic | num.logical) %>%
                  select(code, description, is_numeric) %>%
                  unique() %>%
                  mutate(is_numeric = ifelse(is.na(is_numeric), FALSE, is_numeric)) %>%
                  mutate(is_numeric = ifelse(code == "S1Q1E", FALSE, is_numeric),
                         is_numeric = ifelse(code == "WEIGHT", FALSE, is_numeric),
                         is_numeric = ifelse(code == "STRATUM", FALSE, is_numeric),
                         is_numeric = ifelse(code == "PSU", FALSE, is_numeric),
                         is_numeric = ifelse(code == grepl("scale", 
                                                           description, 
                                                           ignore.case = T), FALSE, is_numeric)
                         ) %>%
                  select(code, is_numeric)

```


# Create Dummy Variables for Missing Continuous

```{r}

column.is.numeric <- w1w2_datatype$code[which(w1w2_datatype$is_numeric == TRUE)]


w1w2suicide_num <- w1w2suicide %>%
                    select(column.is.numeric)


numeric.tocat <- function(x) {
  
  dummy.unknown <- ifelse(x == 99, "unknown", NA)
  x.no.99 <- ifelse(x == 99, NA, x)
  
  quantiles <- ntile(x.no.99, 3) %>% as.character()
  quantiles <- ifelse(quantiles == "1", "low", 
          ifelse(quantiles == "2", "mid", "high"))
  
  quantiles <- ifelse(is.na(quantiles), dummy.unknown, quantiles)
  
  return(quantiles)

}


w1w2_numeric_quantiles <- w1w2suicide_num %>% 
                        map_dfc(~numeric.tocat(.))

names(w1w2_numeric_quantiles)[-1] <- paste(names(w1w2_numeric_quantiles)[-1], "quantiles", sep="_")
w1w2_numeric_quantiles$IDNUM <- w1w2suicide_num$IDNUM

```


# Retype categorical variables

```{r}

mutate.columns.tofactor <- w1w2_datatype$code[which(w1w2_datatype$is_numeric == FALSE)]

w1w2suicide <- w1w2suicide %>% 
                    select(IDNUM, mutate.columns.tofactor) %>%
                    left_join(w1w2_numeric_quantiles, by = "IDNUM") %>%
                    mutate_all(as.factor) %>%
                    mutate_all(fct_explicit_na)
                    

names(w1w2suicide)[-1] <- paste0(names(w1w2suicide)[-1], "_")

write_csv(w1w2_datatype, here("Data","Processed_Data", "NESARC_Codebook_DataType_Variables.csv"))

```


# Check for variables with unique values.

```{r}

remove.unique.cat <- w1w2suicide %>%
                      future_map_dfc(~length(unique(.))) %>%
                      t() %>%
                      as_tibble(rownames = "variable") %>%
                      rename(unique = V1) %>%
                      filter(unique == 1) %>%
                      .$variable

```

# Get Rid of Duplicated Variables

```{r}

duplicated.names <- names(w1w2suicide)[duplicated(as.list(w1w2suicide))]

duplicated.names %>%
    as.data.frame() %>%
    write_csv(x = . , here("Data","Processed_Data", "NESARC_duplicated_Variables.csv"))


w1w2suicide.raw <- w1w2suicide %>%
                    select(-suicide_334_)

w1w2suicide <- w1w2suicide %>%
                select(-duplicated.names)

```

# Write out data

```{r}

rm(w1w2_numeric_quantiles, w1w2suicide_num)

save.image(here("Data","Processed_Data", "w1w2suicidedata_Clean_May2020.RData"))

```

